### 2009-07-30 SC09/abstract.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

% here's the 150 word limited version we submitted to SC
Parallel applications running across thousands of processors must protect
themselves from inevitable system failures. Many applications insulate
themselves from failures by checkpointing. For many applications, checkpointing
into a shared single file is most convenient. With such an approach, the size
of writes are often small and not aligned with file system boundaries.
Unfortunately for these applications, this preferred data layout results in
pathologically poor performance from the underlying file system which is
optimized for large, aligned writes to non-shared files. To address this
fundamental mismatch, we have developed a virtual parallel log structured file
system, PLFS. PLFS remaps an application's preferred data layout into one which
is optimized for the underlying file system. Through testing on PanFS, Lustre,
and GPFS, we have seen that this layer of indirection and reorganization can
reduce checkpoint time by an order of magnitude for several important
benchmarks and real applications.  
} 
}

>>>>

% here's the 150 word limited version we submitted to SC
Parallel applications running across thousands of processors must protect
themselves from inevitable system failures. Many applications insulate
themselves from failures by checkpointing. For many applications, checkpointing
into a shared single file is most convenient. With such an approach, the size
of writes are often small and not aligned with file system boundaries.
Unfortunately for these applications, this preferred data layout results in
pathologically poor performance from the underlying file system which is
optimized for large, aligned writes to non-shared files. To address this
fundamental mismatch, we have developed a virtual parallel log structured file
system, PLFS. PLFS remaps an application's preferred data layout into one which
is optimized for the underlying file system. Through testing on PanFS, Lustre,
and GPFS, we have seen that this layer of indirection and reorganization can
reduce checkpoint time by an order of magnitude for several important
benchmarks and real applications without any application modification.  
} 
}

### 2009-07-30 SC09/arch.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

One difficultly in constructing this global index stems from concurrent
processes that may write the same offset at the same time.  These processes
cannot know which will be the ultimate writer, so they need to synchronize to
determine the appropriate order. This synchronization needs to be exposed to
the file system to be effective~\cite{Gibson95thescotch}. If a logical
clock~\cite{lamport1978} was associated with these synchronizations, its value
could be written to index files so that the merge process could correctly
determine write ordering. Since parallel applications synchronize with a
barrier called by all processes, a simple count of synchronization calls could
be sufficient. In practice checkpoints do not experience overlapping writes, so
at this time \plfs\ has not implemented an overlap conflict resolution scheme.

@@

One interesting nuance is that \plfs\ has a data file for every process but
only a single index file per compute node shared by all processes on that node.
Sharing an index file is easy; by the time \plfs\ sees writes, they have been
merged by the operating system into a single memory space. The operating
systems sends all writes to a single \plfs\ process which ensures index records
are correctly, chronologically appended.  Having a single index greatly reduces
the number of files in a container since current LANL applications run up to
sixteen processes on a node; on the next LANL supercomputer, this could be up
to sixty-four. We tried reducing the number of data files in the same manner.
Write bandwidth was not affected, but reads were slowed for \Term{uniform
restarts} in which reading processes access the file in the same access pattern
as it was written.  The pattern of a single reader accessing a single data file
sequentially lends itself very well to prefetching.  However, due to timing
differences between the write and read phases, multiple processes in a uniform
restart may not always read from a shared file in sequential order.

@@

Because \plfs\ leverages the \upfs\ as much as possible, we give the container
the same logical name as the \plfs\ file. This allows \plfs\ to pass a
\syscall{readdir} system call directly to the \upfs\ and return its result to
the application without any translation.  \plfs\ also handles \syscall{mkdir}
system calls in the same way (\ie\ without any translation). The implication of
leveraging the \upfs\ in this way is that \plfs\ requires some other mechanism
by which to distinguish between regular directories and containers in order to
implement the \syscall{stat} system call correctly. As the \textit{SUID} bit is
rarely used on directories and yet is allowed to be set by the underlying file
system, \plfs\ sets this bit on containers; this does mean, however, that
\plfs\ must disallow setting this bit on a regular directory.

@@

Computing these sums and maximums with every \syscall{stat} call on a \plfs\
file is expensive. Our strategy for speeding this up is to cache recently
computed values in the \Term{metadata subdirectory}. To make this cache as
effective as possible we have each FUSE process cache into this metadata
subdirectory any information it has in its in memory data structures when the
last writer on that node closes the file. On this close, the
FUSE process creates a file named H.L.B.T, where H is the node's hostname, L is
the last offset recorded on that node, B is the sum of the capacity of all
files that this node manages, and T is the maximum timestamp among these files.

When no process has this container open for write, a stat call on the container
is implemented by issuing a \syscall{readdir} on the metadata subdirectory,
then reporting the maximum of the Ls for last byte offset, the maximum of the
Ts for modification timestamp and the sum of the Bs for capacity.

>>>>

One difficulty in constructing the global index stems from multiple
processes that may write the same offset concurrently.  These processes
cannot know which will be the ultimate writer, so they need to synchronize to
determine the appropriate order. This synchronization needs to be exposed to
the file system to be effective~\cite{Gibson95thescotch}. If a logical
clock~\cite{lamport1978} was associated with these synchronizations, its value
could be written to index files so that the merge process could correctly
determine write ordering. Since parallel applications synchronize with a
barrier called by all processes, a simple count of synchronization calls could
be sufficient. In practice checkpoints do not experience overlapping writes, so
at this time \plfs\ has not implemented overlap resolution.

@@

One interesting nuance is that \plfs\ has a data file for every process but
only a single index file per compute node shared by all processes on that node.
Sharing an index file is easy; by the time \plfs\ sees writes, they have been
merged by the operating system into a single memory space. The operating
systems sends all writes to a single \plfs\ process which ensures index records
are correctly, chronologically appended.  Having a single index greatly reduces
the number of files in a container since current LANL applications run up to
sixteen processes on a node; on the next LANL supercomputer, this could be up
to sixty-four. We tried reducing the number of data files similarly.
Write bandwidth was not affected, but reads were slowed for \Term{uniform
restarts} in which reading processes access the file in the same access pattern
as it was written.  The pattern of a single reader accessing a single data file
sequentially lends itself very well to prefetching.  However, due to timing
differences between the write and read phases, multiple processes in a uniform
restart may not always read from a shared file sequentially.

@@

Because \plfs\ leverages the \upfs\ as much as possible, we give the container
the same logical name as the \plfs\ file. This allows \plfs\ to pass a
\syscall{readdir} system call directly to the \upfs\ and return its result
without any translation.  \plfs\ also handles \syscall{mkdir} system calls
without any translation.  Leveraging the \upfs\ in this way requires that
\plfs\ use some other mechanism to distinguish between regular directories and
containers in order to implement the \syscall{stat} system call correctly. As
the \textit{SUID} bit is rarely used on directories and yet is allowed to be
set by the underlying file system, \plfs\ sets this bit on containers; the
implication however is that \plfs\ must disallow setting this bit on a regular
directory.

@@

Computing these sums and maximums with every \syscall{stat} call on a \plfs\
file is expensive. Our strategy for quickening this is to cache recently
computed values in the \Term{metadata subdirectory}. To make this cache as
effective as possible we have each \plfs\ server cache into this metadata
subdirectory any information it has in its memory data structures when the last
writer on that node closes the file. On this close, the FUSE process creates a
file named \Term{H.L.B.T}, where \Term{H} is the node's hostname, \Term{L} is
the last offset recorded on that node, \Term{B} is the sum of the capacity of
all files that this node manages, and \Term{T} is the maximum timestamp among
these files.

When no process has this container open for write, a stat call on the container
issues a \syscall{readdir} on the metadata subdirectory,
then reports the maximum of the offsets for last byte offset, the maximum of
the timestamps for modification timestamp and the sum of the blocks for
capacity.

### 2009-07-30 SC09/eval.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

We present the results of our experimental evaluation in Figure~\ref{fig-eval}.
Eleven of these twelve graphs present one experiment each. The twelfth,
Figure~\ref{eval:summary}, presents a summary. In the majority of these graphs,
the write bandwidth is shown on the \yaxis\ in \MBs\ as a function of the
number of processes. We will note it in the text for those few graphs for
which we deviate from this general configuration. The write bandwidth that we
report is whatever is reported by the particular benchmark; whenever possible,
we report the most conservative value (\ie\ we include open and close times in
our write times, and we either barrier after close or we use the time reported
by the slowest writer). Finally we have attempted to run multiple iterations
for each experiment; where applicable, the standard deviation is therefore
included.

@@

These graphs illustrate how the performance discrepancy between N-N
and N-1 checkpoint patterns is common across PanFS, GPFS,
and Lustre. Remember, as was discussed in Section~\ref{motivation}, switching
to N-N is not a viable option for many applications which are inextricably wed
to an N-1 pattern and are resigned to the attendant loss of bandwidth.
Fortunately, as is evidenced by these graphs, \plfs\ allows these applications
to retain their preferred N-1 pattern while achieving most, if not all, of the
bandwidth available to an N-N pattern. Particularly for the PanFS results,
which were run on our Roadrunner supercomputer, \plfs\ achieves the full
bandwidth of an N-N pattern (\ie\ up to about 31 \GBs). In fact, for several
of the points, an N-1 pattern on \plfs\ actually outperforms an N-N pattern
written directly to PanFS. Although we have yet to fully investigate the exact
reason for this, there are several reasons why this could be the case. The
first is that \plfs\ rearranges writes into a log structured pattern so an N-N
pattern which incurs seeks could do worse than a \plfs\ pattern which appends
only. Secondly, the structure of the \plfs\ container spreads data across
multiple sub-directories within a top-level directory whereas a typical N-N
pattern confines all N files into only a single parent directory.

@@

The second row in Figure~\ref{fig-eval} shows the results of using \plfs\ to
improve the bandwidth of three important LANL applications which use N-1
checkpoint patterns. The application shown in Figure~\ref{eval:lanl1} is the
application whose developers augmented their checkpoint routine with a new
checkpoint library called bulkio, which aggregates and buffers writes into
larger, contiguous writes more friendly to the \upfs. Therefore we compare the
bulkio method writing directly to PanFS to the old MPI-IO method writing
indirectly to PanFS via \plfs.  This data was collected on Roadrunner.  The
next two graphs show similar results; using \plfs\ to rearrange an N-1 pattern
yields significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a
small \rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS
storage system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.
Note that these are extremely important applications at LANL; it has been
estimated that they account for more than half of the total computing cycles at
LANL.  Although these applications must remain anonymous, traces of their IO
are available~\cite{plfs-maps}.

@@

\input{overhead-fig}

@@

\subsubsection{Read Bandwidth}
\input{read-graphs.tex}

@@

We measured the time to read these files using three different read
scenarios. One emulates a \Term{uniform restart} in which the number of
processes resuming the application is the same as the number that wrote the
last checkpoint. In this case, each process within the parallel job reads the
same data in the same order as one of the previous writers. In
contrast, a \Term{non-uniform restart} is one in which the number of
readers is different from the number of writers and the read offsets are
therefore not aligned with the previous writes. In this case, we emulate the
scenario where an application resumes after a failure by simply running on the
newly reduced number of nodes. The number of reads is not affected by the size
of the job for typical N-1 checkpoints. Each read is extracting a region for a
particular variable within the simulation. The size of this region depends on
the number of processes within the job. Each gets $1/Nth$ of the region.
Specifically, for a non-uniform restart in which the number of processes, $N$,
has been reduced by $M$, the size of each read will be $N/(N-M)$ times the size
of the original writes. The third read scenario we emulated is a typical
scenario in which a relatively, small fixed number of processes reads a
checkpoint in order to save it onto an archive system. For these measurements,
we used LANL's archive copy utility with fours readers each running on their
own node; each reader read just one contiguous 5~\GB\ region by issuing 
sequential 1~\MB\ reads.

>>>>

We present the results of our experimental evaluation in Figure~\ref{fig-eval}.
Eleven of these twelve graphs present one experiment each. The twelfth,
Figure~\ref{eval:summary}, is a summary. In the majority of these graphs, the
write bandwidth is shown on the \yaxis\ in \MBs\ as a function of the number of
processes. We will note it in the text for those few graphs for which we
deviate from this general configuration. The write bandwidth that we report is
whatever is reported by the particular benchmark; whenever possible, we report
the most conservative value (\ie\ we include open and close times in our write
times, and we either barrier after close or we use the time reported by the
slowest writer). Benchmarks were run on LANL's Roadrunner supercomputer using
the default parameters unless indicated otherwise.  Finally we have attempted
to run multiple iterations for each experiment; where applicable, the standard
deviation is included.

@@

These graphs illustrate how the performance discrepancy between N-N
and N-1 checkpoint patterns is common across PanFS, GPFS,
and Lustre. Remember, as was discussed in Section~\ref{motivation}, switching
to N-N is not a viable option for many applications which are inextricably wed
to an N-1 pattern and are resigned to the attendant loss of bandwidth.
Fortunately, as is evidenced by these graphs, \plfs\ allows these applications
to retain their preferred N-1 pattern while achieving most, if not all, of the
bandwidth available to an N-N pattern. Particularly for the PanFS results,
which were run on our Roadrunner supercomputer, \plfs\ achieves the full
bandwidth of an N-N pattern (\ie\ up to about 31 \GBs). In fact, for several
of the points, an N-1 pattern on \plfs\ actually outperforms an N-N pattern
written directly to PanFS. Although we have yet to fully investigate the exact
reason for this, there are several reasons why this could be the case. The
first is that \plfs\ rearranges writes into a log structured pattern so an N-N
pattern which incurs seeks could do worse than a \plfs\ pattern which appends
only. Secondly, the structure of the \plfs\ container spreads data across
multiple sub-directories whereas a typical N-N
pattern confines all N files into only a single parent directory.

@@

The second row in Figure~\ref{fig-eval} shows the results of using \plfs\ to
improve the bandwidth of three important LANL applications which use N-1
checkpoint patterns. The application shown in Figure~\ref{eval:lanl1} is the
application whose developers augmented their checkpoint routine with a new
checkpoint library called bulkio, which aggregates and buffers writes into
larger, contiguous writes more friendly to the \upfs. Therefore we compare the
bulkio method writing directly to PanFS to the MPI-IO method writing indirectly
to PanFS via \plfs.  This data was collected on Roadrunner.  The next two
graphs show similar results; using \plfs\ to rearrange an N-1 pattern yields
significantly improved bandwidths.  Figure~\ref{eval:lanl2} was run on a small
\rrz\ which consists of 128 8-core compute nodes with a 66 blade PanFS storage
system, whereas Figure~\ref{eval:lanl3} was run on Roadrunner itself.  Note
that these are extremely important applications at LANL, consuming more than
half of LANL's total cycles.  Although they must remain anonymous, traces of
their IO are available~\cite{plfs-maps}.

@@

\input{overhead-fig}
\input{read-graphs.tex}

@@

\subsubsection{Read Bandwidth}

@@

We measured the time to read these files using three different read scenarios.
One emulates a \Term{uniform restart} in which the number of processes resuming
the application is the same as the number that wrote the last checkpoint. In
this case, each process reads the same data in the same order as one of the
previous writers. In contrast, a \Term{non-uniform restart} is one in which the
number of readers is different from the number of writers and the read offsets
are therefore not aligned with the previous writes. In this case, we emulate
the scenario where an application resumes after a failure by simply running on
the newly reduced number of nodes. The {\em{number}} of reads is not affected
by the size of the job for typical N-1 checkpoints since each read is
extracting a region for a particular variable within the simulation. However,
the {\em{size}} of this region depends on the number of processes within the
job since each gets $1/Nth$ of the region.  Specifically, for a non-uniform
restart in which the number of processes, $N$, has been reduced by $M$, the
size of each read will be $N/(N-M)$ times the size of the original writes. The
third scenario emulated is where a relatively, small fixed number of processes
reads a checkpoint in order to save it onto an archive system. For these
measurements, we used LANL's archive copy utility with fours readers each
running on their own node; each reader reads a single contiguous 5~\GB\ region
by issuing sequential 1~\MB\ reads.

### 2009-07-30 SC09/future.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently an optional mount point on Roadrunner where several
applications have begun using it to improve their N-1 checkpoint performance.

@@

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Many users find it most
convenient to stage their input data sets and their executable code in the same
directory where they write their checkpoint files.  Since \plfs\ is
specifically a checkpoint file system and is not a general purpose file system,
this usage pattern may suffer a performance hit.  These users should instead
stage their input data sets and their executable code in a general purpose file
system and merely direct their N-1 checkpoint files to a \plfs\ directory.

>>>>

\scribble{I removed the 'cute' part of this. Sorry, john}
We initially developed \plfs\ in order to test our hypothesis that an
interposition layer could rearrange checkpoint patterns such that the
convenience of N-1 patterns could be preserved while achieving the bandwidth of
N-N.  However, after achieving such large improvements with real LANL
applications, we were compelled to harden it 
into a production file system. Consisting of about three thousand lines of
code, \plfs\ is currently an optional mount point on Roadrunner where several
applications have begun using it to improve their N-1 checkpoint performance. 
\plfs\ is publically available at
http://sourceforge.net/projects/plfs.

@@

Although \plfs\ works transparently with unmodified applications, in practice
some applications may need to make minor changes.  Many users find it most
convenient to stage their input data sets and their executable code in the same
directory where they write their checkpoint files.  Since \plfs\ is
specifically a checkpoint file system and is not a general purpose file system,
this usage pattern may suffer a performance hit.  These users should instead
stage their input data sets and their executable code in a general purpose file
system and merely direct their N-1 checkpoint files to \plfs.

### 2009-07-30 SC09/intro.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

\scribble{Need to define "its class" --Milo}
% computers are large, they have failures
In June 2008, Los Alamos National Labs (LANL), in partnership with IBM, broke
the petaflop barrier and claimed the top spot on the Top 500 list~\cite{top500}
with the Roadrunner supercomputer~\cite{roadrunner}.  Due to its unique hybrid
architecture, Roadrunner has only 3060 nodes, which is a relatively small
number for a supercomputer of its class.  Unfortunately, even at this size,
component failures are a frequent event.  For large parallel simulations at
LANL, and other High Performance Computing (HPC) sites, which can run for
several months, restarting a calculation from the beginning is not viable,
since the mean time to interrupt for today's supercomputers is measured in days
or hours and not months~\cite{google-failure,schroeder-failure}.

@@

% checkpoint is defense against failure, drives parallel file systems, is
% challengin
Typically these applications protect themselves against failure by periodically
\Term{checkpointing} their progress by saving the state of the application to
persistent storage. After a failure the application can then restart from the most recent 
checkpoint.  Due to the difficulty of reconstructing a consistent
image from multiple asynchronous checkpoints~\cite{lamport1978}, HPC
applications checkpoint synchronously (\ie\ following a barrier).  Synchronous
checkpointing, however, does not eliminate the complexity; it merely shifts it to
the parallel file system which now must coordinate simultaneous access from
thousands of compute nodes.  Even using optimal checkpoint frequencies~\cite{daly-optimal}, 
checkpointing has become the driving workload for
parallel file systems and the challenge it imposes grows with each successively
larger supercomputer~\cite{pcl:99:me,schroeder2007}.  The difficulty of this
challenge can vary greatly depending on the particular pattern of checkpointing
chosen by the application.

% our approach
In this paper, we describe different checkpointing patterns, and show how some
result in very poor storage performance on three of the major HPC
parallel file systems: PanFS, GPFS, and Lustre.  We then posit that an
interposition layer inserted into the existing storage stack can rearrange this
problematic access pattern to achieve much better performance from the
\upfs.  To test this hypothesis, we have developed \plfs, a Parallel
Log-structured File System, which is one such interposition layer.  We present measurements using \plfs\ on several synthetic benchmarks and real applications at multiple HPC supercomputing centers.  The
results confirm our hypothesis: writing to the \upfs\ through
\plfs\ improves checkpoint bandwidth for all tested applications and
benchmarks and on all three studied parallel file systems; in some cases,
bandwidth is raised by several orders of magnitude.

As we shall discuss, and is summarized in Figure~\ref{fig-page1}, PLFS is
already showing very large speed ups for widely used HPC benchmarks and
important real HPC codes, and at extreme scale.

>>>>

\scribble{Need to define "its class" --Milo}
% computers are large, they have failures
In June 2008, Los Alamos National Labs (LANL), in partnership with IBM, broke
the petaflop barrier and claimed the top spot on the Top 500 list~\cite{top500}
with the Roadrunner supercomputer~\cite{roadrunner}.  Due to its unique hybrid
architecture, Roadrunner has only 3060 nodes, which is a relatively small
number for a supercomputer of its class.  Unfortunately, even at this size,
component failures are frequent.  For large parallel simulations at
LANL, and other High Performance Computing (HPC) sites, which can run for
several months, restarting a calculation from the beginning is not viable,
since the mean time to interrupt for today's supercomputers is measured in days
or hours and not months~\cite{google-failure,schroeder-failure}.

@@

% checkpoint is defense against failure, drives parallel file systems, is
% challengin
Typically these applications protect themselves against failure by periodically
\Term{checkpointing} their progress by saving the state of the application to
persistent storage. After a failure the application can then restart from the most recent 
checkpoint.  Due to the difficulty of reconstructing a consistent
image from multiple asynchronous checkpoints~\cite{lamport1978}, HPC
applications checkpoint synchronously (\ie\ following a barrier).  Synchronous
checkpointing, however, does not eliminate the complexity; it merely shifts it to
the parallel file system which now must coordinate simultaneous access from
thousands of compute nodes.  Even using optimal checkpoint frequencies~\cite{daly-optimal}, 
checkpointing has become the driving workload for
parallel file systems and the challenge it imposes grows with each successively
larger supercomputer~\cite{pcl:99:me,schroeder2007}.  The difficulty of this
challenge can vary greatly depending on the particular pattern of checkpointing
chosen by the application. 
In this paper, we describe different checkpointing patterns, and show how some
result in very poor storage performance on three of the major HPC
parallel file systems: PanFS, GPFS, and Lustre.

We then posit that an
interposition layer inserted into the existing storage stack can rearrange this
problematic access pattern to achieve much better performance from the
\upfs.  To test this hypothesis, we have developed \plfs, a Parallel
Log-structured File System, one such interposition layer.  We present measurements using \plfs\ on several synthetic benchmarks and real applications at multiple HPC supercomputing centers.  The
results confirm our hypothesis: writing to the \upfs\ through
\plfs\ improves checkpoint bandwidth for all tested applications and
benchmarks and on all three studied parallel file systems; in some cases,
bandwidth is raised by several orders of magnitude.

As we shall discuss, and is summarized in Figure~\ref{fig-page1}, \plfs\ is
already showing very large speed ups for widely used HPC benchmarks and
important real HPC codes, and at extreme scale.  \plfs\ works on unmodified
applications as an optional mount point on the compute nodes and is designed to
augment, and not replace, an existing \upfs.  \plfs\ consists of three thousand
lines of code and is publically available at
http://sourceforge.net/projects/plfs.

### 2009-07-30 SC09/motivation.tex
final version prepared by John Bent.
Milo it's all yours!



git-svn-id: https://plfs.svn.sourceforge.net/svnroot/plfs/doc@10 979fb63a-54b4-4514-8d67-9ab8ea1b59df

% why all applications not N-N
Since N-N checkpointing derives higher bandwidth than N-1, the obvious path to
faster checkpointing is for application developers to rewrite existing N-1
checkpointing applications to do N-N checkpointing instead.  Additionally, all
new applications should be written to take advantage of the higher bandwidth
available to N-N checkpointing.  Although some developers have gone this route,
many continue to prefer an N-1 pattern even though its disadvantages are well
understood.  There are several advantages to N-1 checkpointing that appeal to
parallel application developers.  One, a single file is much easier to manage
and to archive.  Two, N-1 files usually organize data into an application
specific canonical order that commonly aggregates related data together in
contiguous regions, making visualization of intermediate state simple and
efficient.  Additionally, following a failure, a restart on a different number
of compute nodes is easier to code as the checkpoint format is independent of
the number of processes that captured the checkpoint; conversely, gathering the
appropriate regions from multiple files or from multiple regions within a
single file is more complicated.

@@

Essentially these developers have once again shifted complexity to the parallel
file system for their own convenience.  This is not unreasonable; it has long
been the province of computer systems to make computing more convenient for its
users.  Many important applications have made this choice.  Of the twenty-three
applications listed on the Parallel I/O Benchmarks page~\cite{pio-benchmarks},
at least ten have an N-1 pattern; two major applications at LANL use an N-1
checkpointing pattern as do at least two of the eight applications chosen to
run on Roadrunner during its initial stabilization phase.  N-1 checkpointing is
very important to these applications.  For example, at the core of one of
LANL's N-1 applications is a twenty-year old Fortran checkpointing library. 
About a decade ago, in response to a growing clamor about the
limitations of N-1 checkpointing bandwidth, developers for this application
augmented their checkpointing library with fifteen thousand lines of code.
However, instead of changing the application to write an N-N pattern, they
added a prefer to the IO routiones in which interprocess communication is used to aggregate and buffer writes.
Although they did not improve the performance to match that of other 
applications using an N-N checkpoint pattern, 
this effort was considered a success as they did improve the N-1
performance by a factor of two to three. This new checkpointing library,
called \Term{bulkio}, has been maintained over the past decade and ported to
each new successive supercomputer at LANL~\cite{bent-personal-bulk}. 
Furthermore, N-1 patterns continue to be developed anew in many new
applications.  High level data formatting libraries such as Parallel
NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} offer convenience to application
developers who simply describes the logical format of their data and need no
longer consider how that data is physically organized in the file system.  Once
again this convenience merely shifts the complexity to the \upfs\ 
since these libraries use an N-1 pattern.

>>>>

% why all applications not N-N
Since N-N checkpointing derives higher bandwidth than N-1, the obvious path to
faster checkpointing is for application developers to rewrite existing N-1
checkpointing applications to do N-N checkpointing instead.  Additionally, all
new applications should be written to take advantage of the higher bandwidth
available to N-N checkpointing.  Although some developers have gone this route,
many continue to prefer an N-1 pattern even though its disadvantages are well
understood.  There are several advantages to N-1 checkpointing that appeal to
parallel application developers.  One, a single file is much easier to manage
and to archive.  Two, N-1 files usually organize data into an application
specific canonical order that commonly aggregates related data together in
contiguous regions, making visualization of intermediate state simple and
efficient.  Additionally, following a failure, a restart on a different number
of compute nodes is easier to code as the checkpoint format is independent of
the number of processes that wrote the checkpoint; conversely, gathering the
appropriate regions from multiple files or from multiple regions within a
single file is more complicated.

@@

Essentially these developers have once again shifted complexity to the parallel
file system for their own convenience.  This is not unreasonable; it has long
been the province of computer systems to make computing more convenient for its
users.  Many important applications have made this choice.  Of the twenty-three
applications listed on the Parallel I/O Benchmarks page~\cite{pio-benchmarks},
at least ten have an N-1 pattern; two major applications at LANL use an N-1
checkpointing pattern as do at least two of the eight applications chosen to
run on Roadrunner during its initial stabilization phase.  N-1 checkpointing is
very important to these applications.  For example, at the core of one of
LANL's N-1 applications is a twenty-year old Fortran checkpointing library.
About a decade ago, in response to a growing clamor about the limitations of
N-1 checkpointing bandwidth, developers for this application augmented their
checkpointing library with fifteen thousand lines of code.  However, instead of
changing the application to write an N-N pattern, they added a new IO routine in
which interprocess communication is used to aggregate and buffer writes.
Although they did not improve the performance to match that of other
applications using an N-N checkpoint pattern, this effort was considered a
success as they did improve the N-1 performance by a factor of two to three.
This checkpointing library, called \Term{bulkio}, has been maintained over
the past decade and ported to each new successive supercomputer at
LANL~\cite{bent-personal-bulk}.  Furthermore, N-1 patterns continue to be
developed anew in many new applications.  High level data formatting libraries
such as Parallel NetCDF~\cite{pnetcdf} and HDF5~\cite{hdf5} offer convenience
to application developers who simply describes the logical format of their data
and need no longer consider how that data is physically organized in the file
system.  Once again this convenience merely shifts the complexity to the \upfs\
since these libraries use an N-1 pattern.
